### 2021-06-28 040_solution.tex

\section{Fixing the Compilation of the Contract} \label{sec:fix}