### 2021-07-07 001_abstract.tex
Added abstract

Generics allow the use of type placeholders in order
to write parametrized code, that can be instantiated for each
concrete type provided for the placeholders. In many
programming languages, such as Java, they are implemented by
\emph{erasure}\ie replaced by their upper bound type.
Generics are a powerful feature of programming languages that allows one
to write highly reusable code. Namely, this paper shows their use in the
context of smart contracts for blockchain, in order to implement
a contract for \emph{shared entities} (such as a company shared
by its shareholders),
for the Hotmoka blockchain whose contracts are written in Java.
It shows that a special case of shared entities is the validators set of
the blockchain itself. However, this power comes at the risk of a too permissive
typing of the compiled code, where generics have been erased.
This paper shows
that this allows an attack to the validators set and proposes a solution
that forces the compiler to generate more precise type information than
those arising by erasure.